---
title: "Centrality Plots"
author: "Micha Sam Brickman Raredon"
date: "`r Sys.Date()`"

vignette: >
  %\VignetteIndexEntry{Centrality Plots}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}

---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette details how to use the `Centrality` function of `Connectome`.

## Load dependencies
```{r message=F, warning=F}
library(Seurat)
library(SeuratData)
library(Connectome)
library(ggplot2)
library(cowplot)

```

## Load Data
Again, we will use the cross-platform Pancreas data distributed by SeuratData:
```{r message=F, warning=F,results = 'hide'}
InstallData('panc8')
data('panc8')
table(Idents(panc8))
Idents(panc8) <- panc8[['celltype']]
table(Idents(panc8))
```

## Normalize, Scale and make connectome

``` {r  message=F, warning=F, results = 'hide'}
panc8 <- NormalizeData(panc8)
connectome.genes <- union(Connectome::ncomms8866_human$Ligand.ApprovedSymbol,Connectome::ncomms8866_human$Receptor.ApprovedSymbol)
genes <- connectome.genes[connectome.genes %in% rownames(panc8)]
panc8 <- ScaleData(panc8,features = genes)
panc8.con <- CreateConnectome(panc8,species = 'human',min.cells.per.ident = 75,p.values = F,calculate.DOR = F)
```

## Full system overviews
`Centrality` performs a series of analyses within a single integrated function. First, it takes the entire input network and filters the data based on parameters passed to `FilterConnectome`. Then, it subdivides the remaining network based on the `group.by` argument. Within each sub-network, it then performs a graph-theory-based centrality analysis, ranking each producer and receiver within the sub-network based on their cumulative outgoing and incoming edgeweight and their Kleinberg centrality scores. By default, the resulting plots are normalized over total network edgeweight so that the x-axis is fraction contribution or reception of signal within the sub-network.

We can use `Centrality` to immediately look at an unfiltered edgelist.  Here, we use edgeweights from the normalized data slot, which are always positive:
``` {r  message=F, warning=F, fig.width = 20,fig.height = 10}
Centrality(panc8.con,
           modes.include = NULL,
           min.z = NULL,
           weight.attribute = 'weight_norm',
           group.by = 'mode')
```

However, it is generally more informative to limit the analysis to those edges which meet a user-defined statistical threshold:
``` {r  message=F, warning=F,fig.width = 20,fig.height=10}
Centrality(panc8.con,
           modes.include = NULL,
           weight.attribute = 'weight_norm',
           min.z = 0.25,
           min.pct = 0.1)
```

## Focusing on specific signaling families
`Centrality` has built-in functionality to focus on specific signaling families of interest, which can help to gain a sense of which cells may be dominantly communicating via which pathways:
``` {r  message=F, warning=F,fig.align = "center",fig.width = 10,fig.height = 6}
Centrality(panc8.con,
           modes.include = c('VEGF','NOTCH','TGFB','WNT','ANGPT','EGF'),
           weight.attribute = 'weight_sc',
           min.z = 0.25,
           min.pct = 0.1)
```

## Breaking down a signaling family
It is important to note that a single signaling family is generally made up of multiple mechanisms, and that a ligand can hit multiple receptors and vice-versa.  Therefore, it can be useful to perform centrality analysis for every single mechanism (ligand-receptor pair) present in a sub-network. Here, we do this for every mechanism in the 'NOTCH' family:
``` {r  message=F, warning=F,fig.align = "center",fig.width = 10,fig.height = 8}
Centrality(panc8.con,
           modes.include = c('NOTCH'),
           weight.attribute = 'weight_norm',
           min.z = NULL,
           min.pct = 0.1,
           group.by = 'mechanism')
```


If only specific mechanisms are of interest, `Centrality` can be used to directly inquire about networks based on select genes:
``` {r  message=F, warning=F,fig.align = "center",fig.width = 10,fig.height = 5}
temp <- subset(panc8.con,pair %in% c('TGFB1 - TGFBR1','DLL4 - NOTCH3'))
Centrality(temp,
           weight.attribute = 'weight_norm',
           min.pct = 0.1,
           group.by = 'mechanism')
```

Which suggests that the TGFB1-TGFBR1 network is associated with activated_stellate -> endothelial communication, while the DLL4 - NOTCH3 network is associated with endothelial -> quiescent_stellate commmunication.
