---
title: "Basic Workflow"
author: "Micha Sam Brickman Raredon"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Basic Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Connectome is an R package which facilitates ligand-receptor mapping in single-cell data.  It is built to interact directly with Seurat from Satija Lab. This vignette details the initial workflow.

## Data Preparation

`CreateConnectome` uses the active identity slot in a Seurat 3.0 object to define "nodes" for downstream edgelist creation and network analysis. Node parcellation should be done carefully as it will have a non-trivial effect on downstream connectomics. 

For the purposes of demonstration, we will will run the Connectome workflow on the PBMC data distributed by SeuratData:
```{r load data}
library(SeuratData)
InstallData('pbmc3k')
data('pbmc3k')
pbmc3k
```
To briefly cluster and embed:
``` {r cluster}
pbmc <- pbmc3k
pbmc[["percent.mt"]] <- PercentageFeatureSet(pbmc, pattern = "^MT-")
pbmc <- subset(pbmc, subset = nFeature_RNA > 200 & nFeature_RNA < 2500 & percent.mt < 5)
pbmc <- NormalizeData(pbmc)
pbmc <- FindVariableFeatures(pbmc)
pbmc <- ScaleData(pbmc, features = rownames(pbmc))
pbmc <- RunPCA(pbmc, features = VariableFeatures(object = pbmc))
pbmc <- FindNeighbors(pbmc, dims = 1:10)
pbmc <- FindClusters(pbmc, resolution = 0.5)
pbmc <- RunUMAP(pbmc, dims = 1:10)
DimPlot(pbmc, reduction = "umap")
```
## Running CreateConnectome

`CreateConnectome` creates the initial ligand-receptor mapping against the active identity slot of the input Seurat object and yields a long-form edgelist of all possible cell-cell communication ectors, coupled with extensive metadata which allows downstream statistical filtration. This can be run as follows:
```{r}
library(connectome)
pbmc.con <- CreateConnectome(pbmc3k,LR.database = 'fantom5',species = 'human',p.values = F)
```
```{r, echo=FALSE, results='asis'}
knitr::kable(head(pmbc.con, 10))
```
```

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
