---
title: "Differential Connectomics"
author: "Micha Sam Brickman Raredon"
date: "`r Sys.Date()`"

vignette: >
  %\VignetteIndexEntry{Differential Connectomics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}

---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
`Connectome` can be used to identify and visualize major perturbed cell-cell communication pathways in A/B comparisons of complex tissue systems.

## Load dependencies
```{r message=F, warning=F}
library(Seurat)
library(SeuratData)
library(connectome)
library(ggplot2)
library(cowplot)
library(ComplexHeatmap)
```

## Load Data
To demonstrate differential connectomics using `Connectome`, we will use the interferon-stimulated vs. control PBMC data distributed by SeuratData:
```{r message=F, warning=F}
InstallData('ifnb')
data('ifnb')
table(Idents(ifnb))
Idents(ifnb) <- ifnb[['seurat_annotations']]
table(Idents(ifnb))
```

## Scale and make connectomes
Connectomic networks must first be calculated for each tissue system separately.  To do so:
``` {r  message=F, warning=F,results = 'hide'}
# First identify ligands and receptors which have mapped in the dataset:
connectome.genes <- union(connectome::ncomms8866_human$Ligand.ApprovedSymbol,connectome::ncomms8866_human$Receptor.ApprovedSymbol)
genes <- connectome.genes[connectome.genes %in% rownames(ifnb)]
# Split the object by condition:
ifnb.list <- SplitObject(ifnb,split.by = 'stim')
# Normalize, Scale, and create Connectome:
ifnb.con.list <- list()
for (i in 1:length(ifnb.list)){
  ifnb.list[[i]] <- NormalizeData(ifnb.list[[i]])
  ifnb.list[[i]] <- ScaleData(ifnb.list[[i]],features = rownames(ifnb.list[[i]]))
  ifnb.con.list[[i]] <- CreateConnectome(ifnb.list[[i]],species = 'human',p.values = F)
}
names(ifnb.con.list) <- names(ifnb.list)
```

## Make differential connectome
``` {r  message=F, warning=F,results = 'hide'}
diff <- DifferentialConnectome(ifnb.con.list[[1]],ifnb.con.list[[2]])
```
## Initial Scoring Plot
``` {r  message=F, warning=F,fig.align = "center",fig.width = 50,fig.height = 8}
DifferentialScoringPlot(diff,min.score = 2,min.pct = 0.1,infinity.to.max = T)
```

# Differential Circos Plots
Edgeweights here are the perturbation score, which is **always** positive, by design. However, in conenctomic data, there are 4 separate classes of perturbations, as both the ligand and the receptor can either be differentially increased or decreased.  Here, we investigate each type of perturbation in turn:

## Ligand and receptor are both **UP**:
```{r  message=F, warning=F,fig.align = "center",fig.width = 10,fig.height = 8}
diff.up.up <- subset(diff,ligand.norm.lfc > 0 & recept.norm.lfc > 0 )
CircosDiff(diff.up.up,min.score = 2,min.pct = 0.1,lab.cex = 0.4)
```

## Ligand is **UP** and receptor is **DOWN**:
```{r  message=F, warning=F,fig.align = "center",fig.width = 10,fig.height = 8}
diff.up.down <- subset(diff,ligand.norm.lfc > 0 & recept.norm.lfc < 0 )
CircosDiff(diff.up.down,min.score = 2,min.pct = 0.1,lab.cex = 0.4)
```

## Ligand is **DOWN** and receptor is **UP**:
```{r  message=F, warning=F,fig.align = "center",fig.width = 10,fig.height = 8}
diff.down.up <- subset(diff,ligand.norm.lfc < 0 & recept.norm.lfc > 0 )
CircosDiff(diff.down.up,min.score = 2,min.pct = 0.1,lab.cex = 0.4)
```

## Ligand and receptor are both **DOWN**:
```{r  message=F, warning=F,fig.align = "center",fig.width = 10,fig.height = 8}
diff.down.down <- subset(diff,ligand.norm.lfc < 0 & recept.norm.lfc < 0 )
CircosDiff(diff.down.down,min.score = 2,min.pct = 0.1,lab.cex = 0.4)
```

# Just specific cell interactions:
We can also look at perturbed edges within just a specific community of cells:
```{r  message=F, warning=F,fig.align = "center",fig.width = 10,fig.height = 8}
CircosDiff(diff,min.score = 2,min.pct = 0.1,lab.cex = 0.4,
           sources.include = c('pDC','CD8 T','B'),targets.include = c('CD16 Mono','CD14 Mono'))
```

# Differential signaling in a specific cellular 'niche' due to perturbation:
```{r  message=F, warning=F,fig.align = "center",fig.width = 10,fig.height = 8}
CircosDiff(diff,min.score = 2,min.pct = 0.1,lab.cex = 0.4,
           targets.include = c('CD14 Mono'))
```

# Can look at differential signaling hitting a single receptor:
```{r  message=F, warning=F,fig.align = "center",fig.width = 10,fig.height = 8}
CircosDiff(diff,min.pct = 0.1,lab.cex = 0.4,
           targets.include = c('CD14 Mono'),features = c('CCR5'))
```
